\documentclass{ximera}

\input{../preamble.tex}

\outcome{How do you find the derivative of a product?}
\outcome{How do you find the derivative of a quotient?}
\outcome{Use the product rule to calculate derivatives.}
\outcome{Use the quotient rule to calculate derivatives.}

\title{Product and quotient rules}

\begin{document}

\begin{abstract}
  Now we introduce the rules for calculating the derivatives of
  products and quotients of functions.
\end{abstract}
\maketitle

The derivative of $f(x)g(x)$ is \textbf{not} as simple as
$f'(x)g'(x)$. Never fear, we have a rule for exactly this situation.

\begin{theorem}[The Product Rule]
If $f(x)$ and $g(x)$ are differentiable, then
\[
\ddx f(x)g(x) = f(x)g'(x)+f'(x)g(x).
\]
\end{theorem}


\begin{question}
  Compute the derivative of $f(x) = xe^x$ with respect to $x$.  
\begin{prompt}
$\ddx f(x) = $\answer{e^x + x*e^x}
\end{prompt}  
\end{question}

\begin{question}
  Compute the derivative of $f(x) = e^{2x}$ with respect to $x$.  
\begin{hint} 
Recall that 
\[
e^{2x} = e^x \cdot e^x.
\]
\end{hint}
\begin{prompt}
$\ddx f(x) = $\answer{2*e^x}
\end{prompt}  
\end{question}

We'd like to have a formula to compute
\[
\ddx \frac{f(x)}{g(x)}
\]
if we already know $f'(x)$ and $g'(x)$. For this we have the quotient
rule.

\begin{theorem}[The Quotient Rule]
If $f(x)$ and $g(x)$ are differentiable, then
\[
\ddx \frac{f(x)}{g(x)} = \frac{f'(x)g(x)-f(x)g'(x)}{g(x)^2}.
\]
\end{theorem}

\begin{question}
  Compute the derivative of $f(x) = \frac{1}{x}$ with respect to $x$.  
\begin{prompt}
$\ddx f(x) = $\answer{-1/x^2}
\end{prompt}  
\end{question}


\begin{question}
  Compute the derivative of $f(x) = \frac{x}{e^x}$ with respect to $x$.  
\begin{prompt}
$\ddx f(x) = $\answer{(e^x-x*e^x)/(e^x)^2}
\end{prompt}  
\end{question}





\begin{question}
Write down at least \textbf{five} questions for this lecture. After
you have your questions, label them as ``Level 1,'' ``Level 2,'' or ``Level 3'' where:
\begin{description}
\item[Level 1] Means you know the answer, or know exactly how to do this problem.
\item[Level 2] Means you think you know how to do the problem, or will soon learn how to do the problem.
\item[Level 3] Means you have no idea how to do the problem. 
\end{description}
  \begin{freeResponse}
  \end{freeResponse}
\end{question}

\end{document}
