\documentclass{ximera}
%\input{../preamble.tex}

\outcome{Understand a function as a transformation of its input into its output.}

\outcome{Understand that for every value in the domain, there is exactly one value in the range.}

\outcome{Understand when a function is invertible.}

\outcome{Interpret a function from an algebraic, numerical, graphical
  and verbal perspective.}

\title{Review of functions}

\begin{document}

\begin{abstract}
  This activity will get you thinking about some concepts related to
  functions that are relevant to calculus.
\end{abstract}
\maketitle


\begin{question}
  A number $x$, is to a function $f$, is to a number $f(x)=y$ as:
  %% Alternatively we could use: An angle, is to $\sin$, is to a length as:
  \begin{hint}
      The notation for functions can be confusing. The letter $f$ is
      the ``name'' of the function. Technically, $f(x)$ is the output;
      however, to remind students that $f$ is in fact a function, it
      is common to call the function $f(x)$. For more on this,
      see \href{http://www.purplemath.com/modules/fcnnot.htm}{this link}.
  \end{hint}
  \begin{hint}
    If you are confused, see \href{http://en.wikipedia.org/wiki/Function_(mathematics)}{this link}.
  \end{hint}
  \begin{multipleChoice}
    \choice[correct]{URL, is to a web browser, is to web page.}
    \choice{URL, is to web page, is to a web browser.}
    \choice{Web page, is to a web browser, is to URL.}
    \choice{Web page, is to URL, is to a web browser.}
    \choice{A web browser, is to web page, is to URL.}
    \choice{A web browser, is to URL, is to web page.}
  \end{multipleChoice}
\end{question}


\begin{question}
  On which interval(s) is the function depicted below invertible?
  \begin{image}
    \begin{tikzpicture}
        \begin{axis}[
            width=6in,
            xmin=-5,
            xmax=5,
            ymin=-5,
            ymax=5,
            axis lines =middle,
            every axis y label/.style={at=(current axis.above origin),anchor=south},
            every axis x label/.style={at=(current axis.right of origin),anchor=west},
            axis lines =middle, xlabel=$x$, ylabel=$y$,
            xtick={-5,-4,...,5},
            ytick={-5,-4,...,5},
            grid=both,
            grid style={dashed, gray!50},
          ]
          ]
          \addplot [very thick, blue!50!black, smooth] coordinates {
            (-5,-4)
            (-4,-2)
            (-3,-4)
            (-2,-2)
            (-1.-1)
            (0,1)
            (1,2)
            (2,3)
            (3,4)
            (4,2)
            (5,4)
          };
        \end{axis}
    \end{tikzpicture}
  \end{image}
  \begin{hint}
    If you are confused,
    see \href{http://en.wikipedia.org/wiki/Inverse_function#Graph_of_the_inverse}{this link}.
  \end{hint}
  \begin{multipleChoice}
      \choice[correct]{$-2<x<1$}
      \choice{$-4<x<4$}
      \choice{$-3<x<4$}
      \choice{$-4<x<3$}
      \choice{$-5<x<-4$ and $-3<x<3$ and $4<x<5$}
      \choice{$-4<x<-3$ and $3<x<4$}
      \choice{For all real values of $x$}
      \choice{For no real values of $x$}
    \end{multipleChoice}  
\end{question}


\begin{question}
  Which of the following statements is true?
  %% Alternatively we could use $\sqrt{x}$
  \begin{hint}
    If you are confused, see \href{http://en.wikipedia.org/wiki/Inverse_trigonometric_functions}{this link}.
  \end{hint}
  \begin{multipleChoice}
    \choice[correct]{$\sin\left(\sin^{-1}\left(\frac{1}{2}\right)\right)
      = \frac{1}{2}$} 
    \choice{$\sin^{-1}\left(\sin\left(\frac{5\pi}{2}\right)\right) = \frac{5\pi}{2}$}
    \choice{$\sin^{-1}(x)$ is the inverse function of $\sin(x)$}
    \choice{$\sin^{-1}(x) = \frac{1}{\sin(x)}$}
  \end{multipleChoice}  
\end{question}



\begin{question}
  The expression $\log_b(x) = y$ is equivalent to:
  \begin{hint}
    If you are confused, see \href{http://en.wikipedia.org/wiki/Logarithm}{this link}.
  \end{hint}
  \begin{multipleChoice}
    \choice[correct]{$b^y = x$}
    \choice{$b^x = y$}
    \choice{$x^y = b$}
    \choice{$y^x = b$}
    \choice{$x^b = y$}
    \choice{$y^b = x$}
  \end{multipleChoice}  
\end{question}

\begin{question}
Write down at least \textbf{five} questions for this lecture. After
you have your questions, label them as ``Level 1,'' ``Level 2,'' or ``Level 3'' where:
\begin{description}
\item[Level 1] Means you know the answer, or know exactly how to do this problem.
\item[Level 2] Means you think you know how to do the problem, or will soon learn how to do the problem.
\item[Level 3] Means you have no idea how to do the problem. 
\end{description}
\begin{freeResponse}
  Answers will vary.
\end{freeResponse}
\end{question}

\end{document}
