\documentclass{ximera}

\outcome{Understand the difference between displacement and distance traveled.}
\outcome{Understand the relationship between position, velocity, and acceleration.}
\outcome{Understand how the net change and future value of a function are related to that function's derivative.}
\outcome{Given a velocity function, calculate displacement and distance traveled.}
\outcome{Given a velocity function, find the position function.}
\outcome{Given an acceleration function, find the velocity function.}
\outcome{Calculate net change and future value.}

\title{velocity and net change}

\begin{document}

\begin{abstract}
  The abstract should be a one sentence summary that states the main point of the activity.
\end{abstract}

\maketitle

\begin{question}
  What is the correct answer to this question?

  
    \begin{multipleChoice}
      \choice[correct]{Correct answer}
      \choice{First Distractor}
      \choice{Second Distractor}
      \choice{Third Distractor}
    \end{multipleChoice}  
  
\end{question}

Write down at least \textbf{five} questions for this lecture. After
you have your questions, label them as ``Level 1,'' ``Level 2,'' or ``Level 3'' where:
\begin{description}
\item[Level 1] Means you know the answer, or know exactly how to do this problem.
\item[Level 2] Means you think you know how to do the problem, or will soon learn how to do the problem.
\item[Level 3] Means you have no idea how to do the problem. 
\end{description}
\begin{question}
  \begin{freeResponse}
  \end{freeResponse}
\end{question}

\end{document}
