\documentclass{ximera}

\outcome{What are applications of the derivative?}
\outcome{What can the first and second derivatives of a position function tell us?}
\outcome{What does the derivative tell us about growth?}
\outcome{What does the derivative tell us about cost functions?}
\outcome{Find velocity and acceleration and use these to determine information about position.}
\outcome{Determine average and instantaneous growth rates.}
\outcome{Calculate average and marginal costs.}

\title{derivatives as rates of change}

\begin{document}

\begin{abstract}
  Let's see some basic applications of the derivative.
\end{abstract}
\maketitle

In physical applications, we are often concerned about
\textit{position}, \textit{velocity}, \textit{speed},
\textit{acceleration}.
\begin{align*}
p(t) &= \text{position with respect to time.}\\
v(t) &= p'(t) = \text{velocity with respect to time.}\\
s(t) &= |v(t)| = \text{speed, the absolute value of velocity.}\\
a(t) &=v'(t) = \text{acceleration with respect to time.}
\end{align*}

\paragraph{An example}


The Mostar bridge in Bosnia is $25$ meters above the river
Neretva. For fun, you decided to dive off the bridge. Your position
$t$ seconds after jumping off is
\[
p(t) = -4.9t^2 + 25.
\]

\begin{question}
  When do you hit the water? 
  \begin{prompt}
    Rounding to the nearest hundredth of a second, at around
    \answer{2.26} seconds after your jump.
  \end{prompt}
\end{question}

\begin{question}
  What is your instantaneous velocity as you enter the water?
  \begin{prompt}
    Rounding to the nearest integer, we find \answer{-22} meters per second. 
  \end{prompt}
\end{question}


\begin{question}
  What is your average velocity during your jump?
  \begin{prompt}
    Rounding to the nearest integer, we find \answer{-11} meters per second. 
  \end{prompt}
\end{question}


\begin{question}
Write down at least \textbf{five} questions for this lecture. After
you have your questions, label them as ``Level 1,'' ``Level 2,'' or ``Level 3'' where:
\begin{description}
\item[Level 1] Means you know the answer, or know exactly how to do this problem.
\item[Level 2] Means you think you know how to do the problem, or will soon learn how to do the problem.
\item[Level 3] Means you have no idea how to do the problem. 
\end{description}
  \begin{freeResponse}
  \end{freeResponse}
\end{question}

\end{document}
